---
title: "Homework 2"
output: html_document
---


This is the Rmarkdown file for homework week 2.
Camilla Brunet, Francesca Frisoni, Elisa Perinot.

### Loaded libraries here : rethinking

```{r,library, eval=T, include=FALSE}

library(rethinking)
```

### 1ST PROBLEM
From the Howell1 dataset, consider only the people younger than 13 years
old. Estimate the causal association between age and weight. Assume that
age influences weight through two paths. First, age influences height, and
height influences weight. Second, age directly influences weight through age-
related changes in muscle growth and body proportions.
Draw the DAG that represents these causal relationships. And then write
a generative simulation that takes age as an input and simulates height and
weight, obeying the relationships in the DAG.

```{r,1problem1, eval=T}
data("Howell1")
# filter on age
d<- Howell1[Howell1$age<13,]

# plot distribution of weight
dens(d$weight)

# pag 82
mean(d$weight) # 14 kg

# priors for weight
sample_mu <- rnorm(1e4, 20, 15)
sample_sigma <- runif(1e4, 0, 15)
prior_w <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior_w)
  
# Remove negative weights
prior_w_pos <- prior_w[prior_w > 0]
dens(prior_w_pos)




```

\  

#### SOLUTION

```{r,1problemsol, eval=T}

# optimal solution by Perplexity
# Load data and subset to children under 13
library(rethinking)
data(Howell1)
d <- subset(Howell1, age < 13)

# Fit models to estimate parameters
height_model <- lm(height ~ age, data = d)
weight_model <- lm(weight ~ height + age, data = d)

# Extract coefficients and residual SDs
alpha_h <- coef(height_model)[1]  # Intercept for height
beta_ah <- coef(height_model)[2]  # Slope (age → height)
sigma_h <- sd(residuals(height_model))

alpha_w <- coef(weight_model)[1]  # Intercept for weight
beta_hw <- coef(weight_model)[2]  # Slope (height → weight)
beta_aw <- coef(weight_model)[3]  # Slope (age → weight, direct)
sigma_w <- sd(residuals(weight_model))

# Simulation function
simulate_child <- function(age) {
  # Simulate height from age
  mu_height <- alpha_h + beta_ah * age
  height <- rnorm(length(age), mu_height, sigma_h)
  
  # Simulate weight from height and age
  mu_weight <- alpha_w + beta_hw * height + beta_aw * age
  weight <- rnorm(length(age), mu_weight, sigma_w)
  
  return(data.frame(age = age, height = height, weight = weight))
}

# Example usage: Simulate for ages 5, 8, 11
simulate_child(c(5, 8, 11))



```
WELL DONE!



### 2ND PROBLEM
Estimate the total causal effect of each year of growth on weight.

```{r,2problem, eval=T}


```


### SOLUTION
```{r,2problemsol, eval=T}

```
WELL DONE!



