---
title: "Homework 1"
output: html_document
---

This is the Rmarkdown file for homework week 1.
Camilla Brunet, Francesca Frisoni, Elisa Perinot.

### Loaded libraries here : rethinking

```{r,library, eval=T, include=FALSE}

library(rethinking)
```

### 1ST PROBLEM
Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 
3 water and 11 land. Construct the posterior distribution.


First trial with sequence of 11.
```{r,1problem1, eval=T}
# pag 40
# define grid
p_grid1 <- seq (from=0, to=1, length.out=11)
# define prior
prior1 <- rep(1,11)
# compute likelihood at each value in grid
likelihood1 <- dbinom(3, size=14, prob=p_grid1)
# product of likelihood and prior
unstd.posterior1 <- likelihood1 * prior1
# standardize the posterior, so it sums to 1
posterior1 <- unstd.posterior1 / sum(unstd.posterior1)

# posterior to look at the dataframe
plot(p_grid1, posterior1, type="b", xlab="prob of water", ylab="posterior prob")
mtext("11 points")
```

\  
Second trail with 100.
```{r,1problem2, eval=T}

p_grid2 <- seq (from=0, to=1, length.out=100)
prior2 <- rep(1,100)
likelihood2 <- dbinom(3, size=14, prob=p_grid2)
unstd.posterior2 <- likelihood2 * prior2
posterior2 <- unstd.posterior2 / sum(unstd.posterior2)

plot(p_grid2, posterior2, type="b", xlab="prob of water", ylab="posterior prob")
mtext("100 points")

```

\  
Last trail with 1000.
```{r,1problem3, eval=T}

p_grid3 <- seq (from=0, to=1, length.out=1000)
prior3 <- rep(1,1000)
likelihood3 <- dbinom(3, size=14, prob=p_grid3)
unstd.posterior3 <- likelihood3 * prior3
posterior3 <- unstd.posterior3 / sum(unstd.posterior3)

plot(p_grid3, posterior3, type="b", xlab="prob of water", ylab="posterior prob")
mtext("1000 points")
```

#### SOLUTION

```{r,1problemsol, eval=T}
compute_posterior <- function( W , L , poss=c(0,0.25,0.5,0.75,1) ) {
ways <- sapply( poss , function(q) q^W * (1-q)^L )
post <- ways/sum(ways)
print(data.frame( poss , ways , post=round(post,3) ))
plot(post, type="b", xlab="prob of water", ylab="posterior prob")
mtext("11 points")
}
compute_posterior( 3 , 11 , poss=seq(from=0,to=1,len=11))

```
WELL DONE!



### 2ND PROBLEM
Using the posterior distribution from 1, compute the posterior predictive
distribution for the next 5 tosses of the same globe. I recommend you use
the sampling method.

```{r,2problem, eval=T}
# pag 52
samples3 <- sample(p_grid3, prob=posterior3, size = 10000, replace = T)
plot(samples3)
dens(samples3)

# pag 65-66
w <- rbinom(10000, size=5, prob=samples3)
simplehist(w) #plot(table(w)) other way to plot the w

```


### SOLUTION
```{r,2problemsol, eval=T}
p_samples <- rbeta(1e4,3+1,11+1)
W_sim <- rbinom(1e4,size=5,p=p_samples)
plot(table(W_sim))

```
WELL DONE!



### ADDITIONAL
2E1-2E4 on page 46 of SR2023 - not completed

```{r,1esbook, eval=F}

```

3E1-3E7 pag 69

```{r,2esbook, eval=T}
# distribution given by the book
p_grid <- seq (from=0, to=1, length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(6, size=9, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
set.seed(100)
samples <- sample(p_grid, prob=posterior, size = 10000, replace = T)

# 3e1
# how much post prob under p=0.2
print(quantile(samples, 0.2))

# 3e2
# how much post prob under p=0.8
print(quantile(samples, 0.8))

# 3e3
# how much post prob between p=0.2 and 0.8 
print(quantile(samples, c(0.2,0.8)))

# 3e4
# 20% of the post prob lies below which value of p
print(sum(posterior[p_grid<0.2]))

plot(posterior, type="b")
abline(h=sum(posterior[p_grid<0.2]))

# 3e5
# 20% of the post prob lies above which value of p
print(sum(posterior[p_grid>0.2]))

# 3e6
# which values of p contain the narrowest interval equal to 66% of the post prob
print(HPDI(samples, prob = 0.66))

# 3e7
# not completed


```




